Creating the ``r params$package_name`` R package

### 2025-07-23 create-FSHybridPLS.Rmd
new formula aligns with the old when lambda=0

```{r}
testthat::test_that("rep_fd correctly replicates a list of fd objects", {
  suppressPackageStartupMessages(library(fda))

  # Create a simple basis
  basis <- create.bspline.basis(c(0, 1), 5)

  # Create single-sample fd object
  coef_vec <- 1:5
  fd1 <- fd(coef = matrix(coef_vec, ncol = 1), basisobj = basis)

  # Replicate the list
  rep_list <- rep_fd(list(fd1, fd1), 3)

  # Check: list length unchanged
  testthat::expect_equal(length(rep_list), 2)

  # Check each replicated fd object
  for (fd_obj in rep_list) {
    testthat::expect_s3_class(fd_obj, "fd")
    testthat::expect_equal(dim(fd_obj$coefs), c(5, 3))  # 5 basis × 3 samples
    for (i in 1:3) {
      testthat::expect_equal(fd_obj$coefs[, i], coef_vec)
    }
  }

  # Error if not single-sample
  fd_multi <- fd(coef = matrix(1:10, nrow = 5, ncol = 2), basisobj = basis)
  testthat::expect_error(rep_fd(list(fd_multi)), "Each fd object in the list must have one column")

  # Error on non-fd input
  testthat::expect_error(rep_fd(list(1, "not_fd")), "Input must be a list of 'fd' objects")
})

@@

  gram_list <- vector("list", n_functional)
  n_basis_list <- numeric(n_functional)

@@

    basis_i <- fd_i$basis
    gram_list[[i]] <- compute_gram_matrix(basis_i)
    gram_deriv2_list[[i]] <- compute_gram_matrix_deriv2(basis_i) # New function call

@@

Let's write some tests to make sure the function behaves as desired:
```{r}
testthat::test_that("predictor_hybrid constructor works as expected (S3, updated)", {
  suppressPackageStartupMessages(library(fda))

  # Create scalar matrix
  Z <- matrix(rnorm(100 * 5), nrow = 100, ncol = 5)

  # Create two fd objects with 100 replications
  basis1 <- create.bspline.basis(c(0, 1), nbasis = 10)
  basis2 <- create.bspline.basis(c(0, 1), nbasis = 20)

  coef1 <- matrix(rnorm(10 * 100), nrow = 10)  # 10 basis × 100 samples
  coef2 <- matrix(rnorm(20 * 100), nrow = 20)  # 20 basis × 100 samples

  fd1 <- fd(coef = coef1, basisobj = basis1)
  fd2 <- fd(coef = coef2, basisobj = basis2)

  # Construct object
  obj <- predictor_hybrid(Z = Z, functional_list = list(fd1, fd2))

  # Assertions
  testthat::expect_s3_class(obj, "predictor_hybrid")
  testthat::expect_equal(nrow(obj$Z), 100)
  testthat::expect_equal(ncol(obj$Z), 5)
  testthat::expect_equal(length(obj$functional_list), 2)
  testthat::expect_equal(obj$n_basis_list, c(10, 20))
  testthat::expect_equal(obj$n_sample, 100)
  testthat::expect_equal(obj$n_functional, 2)
  testthat::expect_equal(obj$n_scalar, 5)
  testthat::expect_equal(dim(obj$gram_list[[1]]), c(10, 10))
  testthat::expect_equal(dim(obj$gram_list[[2]]), c(20, 20))
})
```

@@

**unit test**
```{r}
testthat::test_that("hybrid_from_coef reconstructs single-sample predictor_hybrid correctly", {
  suppressPackageStartupMessages(library(fda))

  make_fd <- function(coefs, basis) {
    fd(coef = coefs, basisobj = basis)
  }

  basis <- create.bspline.basis(c(0, 1), nbasis = 4)
  n_sample <- 5
  n_scalar <- 2
  n_functional <- 2

  # Create dummy functional list with zeros
  fd1 <- make_fd(matrix(0, 4, n_sample), basis)
  fd2 <- make_fd(matrix(0, 4, n_sample), basis)
  Z <- matrix(0, n_sample, n_scalar)
  obj_template <- predictor_hybrid(Z = Z, functional_list = list(fd1, fd2))

  # Create coefficient vector for reconstruction: 4 + 4 + 2 = 10
  xi_star <- c(1:10)

  # Apply reconstruction
  obj_single <- predictor_hybrid_from_coef(obj_template, xi_star)

  # Checks
  testthat::expect_s3_class(obj_single, "predictor_hybrid")
  testthat::expect_equal(obj_single$n_sample, 1)
  testthat::expect_equal(obj_single$Z, matrix(c(9, 10), nrow = 1))

  # Check functional coefficients (use coef() not $coefs)
  testthat::expect_equal(drop(coef(obj_single$functional_list[[1]])), 1:4)
  testthat::expect_equal(drop(coef(obj_single$functional_list[[2]])), 5:8)
})
```

@@

Here are the unit tests:
| Case                    | Description                         | Expected Outcome |
| ----------------------- | ----------------------------------- | ---------------- |
| ✅ Same sample size      | Functional and scalar values summed | Pass             |
| ✅ Broadcast one → many  | One-sample input broadcasted        | Pass             |
| ✅ Broadcast many ← one  | One-sample other broadcasted        | Pass             |
| ❌ Basis mismatch        | Different functional basis          | Error            |
| ❌ Irregular sample size | Neither matches nor is singleton    | Error            |
| ❌ Different dimensions  | Functional or scalar count mismatch | Error            |
```{r}
testthat::test_that("add.predictor_hybrid handles broadcasting and basis compatibility", {
  suppressPackageStartupMessages(library(fda))

  # Create shared B-spline basis
  basis <- create.bspline.basis(c(0, 1), nbasis = 5)

  # Helper to generate a valid coefficient matrix
  coef_mat <- function(val, n_sample, basis_obj) {
    nbasis <- basis_obj$nbasis
    matrix(val, nrow = nbasis, ncol = n_sample)
  }

  # Helper to create a list of fd objects (2 functional predictors)
  fd_list <- function(val, n_sample, basis_obj = basis) {
    list(
      fd(coef = coef_mat(val, n_sample, basis_obj), basisobj = basis_obj),
      fd(coef = coef_mat(val, n_sample, basis_obj), basisobj = basis_obj)
    )
  }

  # Scalar predictors
  Z1 <- matrix(1, nrow = 1, ncol = 2)
  Z3 <- matrix(3, nrow = 3, ncol = 2)

  # Functional predictors (same basis)
  F1 <- fd_list(1, 1)
  F3 <- fd_list(3, 3)

  # Create predictor_hybrid objects
  x1 <- predictor_hybrid(Z1, F1)  # 1 sample
  x3 <- predictor_hybrid(Z3, F3)  # 3 samples

  # Broadcast x1 to match x3
  result <- add.predictor_hybrid(x1, x3, alpha = 1)

  # Check metadata
  testthat::expect_s3_class(result, "predictor_hybrid")
  testthat::expect_equal(result$n_sample, 3)
  testthat::expect_equal(result$n_scalar, 2)
  testthat::expect_equal(result$n_functional, 2)

  # Check scalar component
  expected_Z <- matrix(4, nrow = 3, ncol = 2)  # 1 + 3 = 4
  testthat::expect_equal(result$Z, expected_Z)

  # Check functional component
  for (fd_out in result$functional_list) {
    testthat::expect_equal(unname(coef(fd_out)), matrix(4, nrow = 5, ncol = 3))
  }

  # Test symmetry: broadcasting x3 + x1
  result2 <- add.predictor_hybrid(x3, x1, alpha = 1)
  testthat::expect_equal(result2$Z, expected_Z)
  for (fd_out in result2$functional_list) {
    testthat::expect_equal(unname(coef(fd_out)), matrix(4, nrow = 5, ncol = 3))
  }

  # Incompatible sample sizes (should fail)
  x_bad <- predictor_hybrid(matrix(1, 2, 2), fd_list(1, 2))
  testthat::expect_error(add.predictor_hybrid(x3, x_bad), "incompatible for broadcasting")

  # Incompatible functional basis (should fail)
  bad_basis <- create.bspline.basis(c(0, 1), nbasis = 6)
  bad_fd <- fd_list(2, 3, bad_basis)
  x_basis_mismatch <- predictor_hybrid(Z3, bad_fd)
  testthat::expect_error(add.predictor_hybrid(x3, x_basis_mismatch), "same basis")
})

```
### subtr.predictor_hybrid
Performs element-wise subtraction of two `predictor_hybrid` objects.   Internally uses `add.predictor_hybrid(input, other, alpha = -1)` to compute the result by negating the second operand.

@@

Here is a test code with two functional predictors:
```{r}
testthat::test_that("scalar_mul.predictor_hybrid works with two functional predictors", {
  suppressPackageStartupMessages(library(fda))

  # Scalar part (3 observations, 2 scalar predictors)
  Z <- matrix(2, nrow = 3, ncol = 2)

  # Functional part: shared B-spline basis
  basis <- create.bspline.basis(c(0, 1), nbasis = 5)
  fd1 <- fd(coef = matrix(2, nrow = 5, ncol = 3), basisobj = basis)
  fd2 <- fd(coef = matrix(2, nrow = 5, ncol = 3), basisobj = basis)

  # Construct hybrid object using new simplified constructor
  obj <- predictor_hybrid(Z = Z, functional_list = list(fd1, fd2))

  # Apply scalar multiplication
  scaled <- scalar_mul.predictor_hybrid(obj, scalar = 4)

  # Check class
  testthat::expect_s3_class(scaled, "predictor_hybrid")

  # Check scalar predictor was scaled
  testthat::expect_equal(scaled$Z, Z * 4)

  # Check each functional component's coefficients
  expected_coef <- matrix(8, nrow = 5, ncol = 3)
  for (fd_obj in scaled$functional_list) {
    testthat::expect_equal(
      unname(coef(fd_obj)),
      unname(expected_coef)
    )
  }
})
```
### inprod.predictor_hybrid
- The inner product of $f_1=(f_1^{(1)}, \ldots,f_1^{(K)})$ and $f_2=(f_2^{(1)}, \ldots,f_2^{(K)})$ in $\mathcal{F}$ is defined as 
$$\langle f_1, f_2\rangle_\mathcal{F} =  \sum_{k=1}^K \langle f_1^{(k)}, f_2^{(k)}\rangle_{L^2} = \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k,$$ with norm 
$$\Vert f_1 \Vert_\mathcal{F} = \langle f_1,f_1 \rangle_\mathcal{F}^{1/2} = \{ \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k)^2 dt_k\}^{1/2}.$$
- We define the inner product between any two hybrid objects, $\mathbf{h}_1 = (f_1, \mathbf{v}_1)$ and $\mathbf{h}_2 = (f_2, \mathbf{v}_2)$, as
$$
    \langle \mathbf{h}_1, \mathbf{h}_2\rangle_{\mathcal{H}} = \langle f_1, f_2\rangle_\mathcal{F} +  \langle  \mathbf{v_1}, \mathbf{v_2} \rangle = \sum \limits_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k + \sum \limits_{r=1}^p v_{1r}v_{2r}, 
$$
with norm $\Vert \cdot \Vert_\mathcal{H} = \langle \cdot,\cdot \rangle_\mathcal{H}^{1/2}$.

@@

### get_penalty_hybrid
Constructs a matrix defined in \label{def:J_dotdot_ast}
$$
    \ddot{J}^\ast=\mathrm{blkdiag}(\ddot{J}^{(1)}, \cdots, \ddot{J}^{(K)}, 0_{p \times p}) \in \mathbb{R}^{(MK+p) \times (MK+p)},
$$
where
 $\ddot{J}^{(k)}$ is  a $M \times M$ matrix defined in \label{def:dodot_J_k} as the gram matrix formed by the second derivativ of the basis functions
$$
\ddot{J}^{(k)} = \left[ \int_{\mathcal{T}_k} \ddot{b}^{(k)}_m(t) \ddot{b}^{(k)}_n(t) \, dt \right]_{m,n=1}^M.
$$
```{r}
#' Construct the roughness penalty matrix for a predictor_hybrid object
#'
#' Returns a block diagonal matrix representing roughness penalties for each functional predictor
#' and zeros for the scalar predictors (scalar predictors are not penalized). Each functional component uses its own basis penalty matrix, computed via second derivative penalty via `getbasispenalty()` from the `fda` package
#'
#' @param W A `predictor_hybrid` object.
#'
#' @return A block diagonal penalty matrix of dimension `(sum(n_basis_list) + n_scalar)^2`.
#' @export
get_penalty_hybrid <- function(W){
  J_dotdot_star <- fda::getbasispenalty(W$functional_list[[1]]$basis)
  for (ii in 2:(W$n_functional)){
    J_dotdot_star <- Matrix::bdiag(
      J_dotdot_star,
      fda::getbasispenalty(W$functional_list[[ii]]$basis)
    )
    }
  J_dotdot_star <- Matrix::bdiag(
    J_dotdot_star,
    matrix(0, nrow = W$n_scalar, ncol = W$n_scalar)
    )
  return(J_dotdot_star)
  }
```

@@

### helper: get_xi_hat_linear
```{r}
get_xi_hat_linear_new <- function(W, y){
  # new
  ## step 1: inner products
  u <- list()
  u_j <- B %*% t(Theta_j) %*% y
  for (j in 1:K){
    Theta_t <- W$functional_list[[j]]$coefs
    B <- W$gram_list[[j]]
    u[[j]] <- B %*% Theta_t %*% y
  }
  v <- t(W$Z) %*% y

  ## step 2: squared norm
  q <- sum(v^2)
  for (j in 1:K){
    Theta_t <- W$functional_list[[j]]$coefs
    R <- W$gram_list[[j]]
    u[[j]] <- B %*% Theta_t %*% y
  }  
  # old
  n <- W$n_sample
  K <- W$n_functional
  normalizer <- 0

  for (i in 1:K){  ## tPhi_tC : C multiplied by Phi. Scalable to arbitrary K

    d[[i]] <- Theta_t %*% y
    normalizer <- normalizer + t(y) %*% t(Theta_t) %*% B %*% Theta_t %*% y
  }
  normalizer <- normalizer+ t(y) %*% W$Z %*% t(W$Z) %*% y
  d_vec <- do.call(c, d)  # or: unlist(d)
  d_vec <- d_vec/ sqrt(normalizer)
  return(d_vec)
}
```

@@

@@

xi_hat_lin <- get_xi_hat_linear(W, y)
xi_hat_lin_new <- get_xi_hat_linear_new(W, y)
# Wrap into predictor_hybrid object using format template
xi_hat_direct <- predictor_hybrid_from_coef(format = W_template, coef = xi_hat_lin)
norm_val <- sqrt(inprod.predictor_hybrid(xi_hat_direct))
xi_hat_direct <- scalar_mul.predictor_hybrid(xi_hat_direct,  1 / norm_val)
xi_hat_direct_new <- predictor_hybrid_from_coef(format = W_template, coef = xi_hat_lin_new)

>>>>

@@

  gram_list <- gram_deriv2_list <- vector("list", n_functional)
  n_basis_list <- numeric(n_functional)

@@

    basis_i <- fd_i$basis
    gram_list[[i]] <- compute_gram_matrix(basis_i)
    gram_deriv2_list[[i]] <- fda::getbasispenalty(basis_i) # New

@@

@@

@@

### subtr.predictor_hybrid
Performs element-wise subtraction of two `predictor_hybrid` objects.   Internally uses `add.predictor_hybrid(input, other, alpha = -1)` to compute the result by negating the second operand.

@@

### inprod.predictor_hybrid
- The inner product of $f_1=(f_1^{(1)}, \ldots,f_1^{(K)})$ and $f_2=(f_2^{(1)}, \ldots,f_2^{(K)})$ in $\mathcal{F}$ is defined as 
$$\langle f_1, f_2\rangle_\mathcal{F} =  \sum_{k=1}^K \langle f_1^{(k)}, f_2^{(k)}\rangle_{L^2} = \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k,$$ with norm 
$$\Vert f_1 \Vert_\mathcal{F} = \langle f_1,f_1 \rangle_\mathcal{F}^{1/2} = \{ \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k)^2 dt_k\}^{1/2}.$$
- We define the inner product between any two hybrid objects, $\mathbf{h}_1 = (f_1, \mathbf{v}_1)$ and $\mathbf{h}_2 = (f_2, \mathbf{v}_2)$, as
$$
    \langle \mathbf{h}_1, \mathbf{h}_2\rangle_{\mathcal{H}} = \langle f_1, f_2\rangle_\mathcal{F} +  \langle  \mathbf{v_1}, \mathbf{v_2} \rangle = \sum \limits_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k + \sum \limits_{r=1}^p v_{1r}v_{2r}, 
$$
with norm $\Vert \cdot \Vert_\mathcal{H} = \langle \cdot,\cdot \rangle_\mathcal{H}^{1/2}$.

@@

### get_penalty_hybrid
Constructs a matrix defined in \label{def:J_dotdot_ast}
$$
    \ddot{J}^\ast=\mathrm{blkdiag}(\ddot{J}^{(1)}, \cdots, \ddot{J}^{(K)}, 0_{p \times p}) \in \mathbb{R}^{(MK+p) \times (MK+p)},
$$
where
 $\ddot{J}^{(k)}$ is  a $M \times M$ matrix defined in \label{def:dodot_J_k} as the gram matrix formed by the second derivativ of the basis functions
$$
\ddot{J}^{(k)} = \left[ \int_{\mathcal{T}_k} \ddot{b}^{(k)}_m(t) \ddot{b}^{(k)}_n(t) \, dt \right]_{m,n=1}^M.
$$

@@

### helper: get_xi_hat_linear
```{r}
get_xi_hat_linear_pen <- function(W, y, lambda){
  # new
  ## step 1: inner products
    n <- W$n_sample
  K <- W$n_functional
  u <- gamma <- list()

  v <- t(W$Z) %*% y
  q <- sum(v^2)
  for (j in 1:K){
    Theta_t <- W$functional_list[[j]]$coefs
    B <- W$gram_list[[j]]
    u[[j]] <- B %*% Theta_t %*% y

    # linear system
    R <- W$gram_list[[j]] + lambda[j] * W$gram_deriv2_list[[j]]
    U_cholesky <- chol(R)
    L_cholesky <- t(U_cholesky)
    intermediate <- forwardsolve(L_cholesky, u[[j]])
    gamma[[j]] <- backsolve(U_cholesky, intermediate)

     #q
    q <- q + sum( u[[j]] * gamma[[j]] )
  }
  d_vec <- c(do.call(c, gamma), v)  # or: unlist(d)
  ## step 2: squared norm
  d_vec <- d_vec/ sqrt(q)
  return(d_vec)
}
```

@@

# --- Step 4: Compute constraint matrix ---
lambda <- rep(0, W$n_functional)
#constr_mat <- get_constraint_matrix_hybrid(W, lambda)
#L <- chol(constr_mat)

@@

xi_hat_pen <- get_xi_hat_linear_pen(W, y, lambda)
xi_hat_lin_new <- get_xi_hat_linear_new(W, y)
# Wrap into predictor_hybrid object using format template
xi_hat_direct <- predictor_hybrid_from_coef(format = W_template, coef = xi_hat_lin)
norm_val <- sqrt(inprod.predictor_hybrid(xi_hat_direct))
xi_hat_direct <- scalar_mul.predictor_hybrid(xi_hat_direct,  1 / norm_val)
xi_hat_direct_new <- predictor_hybrid_from_coef(format = W_template, coef = xi_hat_lin_new)
